---
title: "Week 6: Day 2 Homework"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 78
---

# Distributions homework

You are given a cut of a database containing details of mobile phones and
their customer reviews on an online shopping site in file 20190928-items.csv.
Each row in the data corresponds to one model of mobile phone. Use the data to
answer the following questions.

As part of the MVP we want you not to just run the code but also have a go at
intepreting the results - this is the main learning objective from this
homework. Don't worry if you aren't sure, it's ok to get it wrong, but give it
a go, through trying we will learn.

# 1 MVP

## Question 1.

Load the data, clean_names() and explore it.

```{r}
library(prob)
library(gtools)
library(tidyverse)
library(janitor)
library(skimr)
```

```{r}
mobile_phones <- read_csv("data/20190928-items.csv")
```

```{r}
head(mobile_phones)
```

```{r}
summarise(mobile_phones)
```

```{r}
glimpse(mobile_phones)
```

## Question 2.

Find the brand with the highest numbers of phones in the dataset.

```{r}
brands <- mobile_phones %>% 
  group_by(brand) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))

brands

# Samsung has the highest number of phones in the dataset
```

## Question 3.

For your top brand, plot the distribution of phone ratings as a probability
density, overlaying a fitted normal distribution. Do you think the normal
distribution provides a good model of these ratings? Hint You will need to
calculate the mean() and sd() of the ratings to plot the appropriate normal
distribution. Alternatively, investigate using the fitdistr() function in the
MASS package

```{r}
# First create a table for just the samsung data
# We only need the brand and the rating column to simplify
samsung <- mobile_phones %>% 
  filter(brand == "Samsung") %>% 
  select(brand, rating)
```

```{r}
# Have a quick look at the stats
samsung_stats <- samsung %>%
  summarise(
    num = n(),
    mean = mean(rating),
    sd = sd(rating)
  )
samsung_stats

# 397 rows
# mean rating of 3.5733
# standard deviation of 0.6847859
```

```{r}
# Plot the distribution of phone ratings as a probability
# density, overlaying a fitted normal distribution
samsung %>% 
  ggplot(aes(x = rating)) +
  geom_histogram(aes(y = ..density..), col = "white", bins = 25) +
  stat_function(
    fun = dnorm, 
    args = list(
      mean = samsung_stats$mean, 
      sd = samsung_stats$sd
    ),
    col = "red"
  )
```

The data is clearly left-skewed, with the tale to the left stretching towards
the outlying data at 1. Perhaps the outlier is creating the skew and the bulk
of the data is showing a normal distribution. The bulk of the ratings lie
between 3 and 4 reflecting the mean of 3.5733.

## Question 4.

We can be more quantitative in comparing the distribution of top brand ratings
with a normal distribution. You calculated the mean() and sd() of your
distribution in an earlier question. Use these values to calculate the
proportions of ratings within one-, two- and three standard deviations of the
mean. Compare these proportions with those you expect for a normal
distribution.

Hint You can use a filter() to get only the ratings within one-, two- and
three standard deviations of the mean, and then think about how to calculate
what proportion each of these filtered sets make of the total number of
ratings. You'll need to use three separate filter() and summarise() pipes to
do this.

Further hint - code to filter within one SD Here's the code to filter and
summarise the proportion of ratings within one standard deviation of the mean.

```{r}
  filter(rating >= samsung_stats$mean - samsung_stats$sd) %>%
  filter(rating <= samsung_stats$mean + samsung_stats$sd) %>%
  summarise(prop_within_1sd = n() / nrow(samsung_ratings))
```

```{r}
# To find the values within one sd we need to filter between 
# Mean - SD and M + SD
prop_within_1sd <- samsung %>% 
  filter(rating >= samsung_stats$mean - samsung_stats$sd) %>%
  filter(rating <= samsung_stats$mean + samsung_stats$sd) %>%
  summarise(prop_within_1sd = n() / nrow(samsung)) 
prop_within_1sd
```

## Comments for one standard deviation

The proportion of the Samsung ratings within one standard deviation is
0.7934509 or 79.34% The normal distribution has 68.27% within one sd - so the
Samsung data is more concentrated.

```{r}
# To find the values within two sd we need to filter between 
# Mean - 2*SD and M + 2*SD
prop_within_2sd <- samsung %>% 
  filter(rating >= samsung_stats$mean - (2*samsung_stats$sd)) %>%
  filter(rating <= samsung_stats$mean + (2*samsung_stats$sd)) %>%
  summarise(prop_within_2sd = n() / nrow(samsung)) 
prop_within_2sd
```
## Comments for two standard deviations

The proportion of the Samsung ratings within two standard deviations is
0.9193955 or 91.94% The normal distribution has 95.45% within one sd - so the
Samsung data is less than we would expect for the normal distribution.

```{r}
# To find the values within three sd we need to filter between 
# Mean - 3*SD and M + 3*SD
prop_within_3sd <- samsung %>% 
  filter(rating >= samsung_stats$mean - (3*samsung_stats$sd)) %>%
  filter(rating <= samsung_stats$mean + (3*samsung_stats$sd)) %>%
  summarise(prop_within_3sd = n() / nrow(samsung)) 
prop_within_3sd
```
## Comments for three standard deviations

The proportion of the Samsung ratings within three standard deviations is
0.9798489 or 97.98% The normal distribution has 95.45% within one sd - so the
Samsung data is more than we would expect for the normal distribution.



