---
title: "R Notebook"
output: html_notebook
---
# Import Files ----
```{r}
library(CodeClanData)
```
```{r}
drinks_content
```

```{r}
drinks_content <- drinks_content
```

# Question One ----

```{r}
# Glimpse the data 
glimpse(drinks_content)
```

```{r}
# Short representation of the data 
# Data is 242 rows and 18 columns
# Data contains mostly numeric and some character values
# Data describes the make up of the Starbucks drinks
str(drinks_content)
```

# Question Two ----

```{r}
# Select category, beverage, prep and calories
#Drop any NAs found in the Calories column
drinks_content %>%
  select (Beverage_category, Beverage, Beverage_prep, Calories)  %>%
  drop_na(Calories)
```

# Question Three ----

```{r}
# Calculate the calorie difference from 135 calories
drinks_content %>%
  mutate(calorie_diff = Calories - 135) 
```

# Question Four ----

```{r}
# Summarise drinks by category 
# Display the top 3 groups
drink_groups <- group_by(drinks_content, Beverage_category)

summarise (drink_groups, Calories = mean(Calories)) %>%
  arrange (desc(Calories)) %>%
  head(3)
```

Smoothies, Frappuccino Blended Coffee and Signature Espresso Drinks are the 
most calorific categories of drinks.

```{r}
# Summarise drinks by category 
# Display the bottom 3 groups
drink_groups <- group_by(drinks_content, Beverage_category)

summarise (drink_groups, Calories = mean(Calories)) %>%
  arrange (desc(Calories)) %>%
  tail(3)
```
Classic Espresso Drinks, Shaken Iced Beverages and Coffee are the lowest
calorie groups.

# Question 5 ----
```{r}
# Summarise drinks by Beverage_prep 
# Display the average number
prep_groups <- group_by(drinks_content, Beverage_prep)

summarise (prep_groups, Calories = mean(Calories)) %>%
  arrange (desc(Calories)) 
```

# Question 6 ----

```{r}
# Find the overall mean for all drinks
mean(drinks_content$Calories)
```

The overall mean is 193.8719
Returning to the code from Question 5 - which are above this average?

```{r}
# Summarise drinks by Beverage_prep 
# Display the average number
prep_groups <- group_by(drinks_content, Beverage_prep)

summarise (prep_groups, Calories = mean(Calories)) %>%
  filter (Calories >= 193.8719) 
```
There are 5 drink preps with above average calories:
  2% Milk
  Grande Nonfat Milk
  Soymilk
  Venti Nonfat Milk
  Whole Milk
  
# Question 7 ----

```{r}
# Summarise drinks by Beverage_prep 
# Display the average number
# Sort by the lowest first
prep_groups <- group_by(drinks_content, Beverage_prep)

summarise (prep_groups, Calories = mean(Calories)) %>%
  arrange (Calories) 
```
The coffee type with the lowest calories is the solo drinks.
Looks like it is the fat that causes all of the problems!
