---
title: "R Notebook"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
library(tidyverse)
library(GGally)
library(ggfortify)
library(janitor)
library(fastDummies)
library(mosaic)
library(mosaicData)
```



# File information

About this file

A data frame with 53940 rows and 10 variables:

price price in US dollars (\$326--\$18,823)

carat weight of the diamond (0.2--5.01)

cut quality of the cut (Fair, Good, Very Good, Premium, Ideal)

colour diamond colour, from J (worst) to D (best)

clarity a measurement of how clear the diamond is (I1 (worst), SI2, SI1,
VS2, VS1, VVS2, VVS1, IF (best))

x length in mm (0--10.74)

y width in mm (0--58.9)

z depth in mm (0--31.8)

depth total depth percentage = z / mean(x, y) = 2 \* z / (x + y)
(43--79)

table width of top of diamond relative to widest point (43--95)

# **1 MVP**

1.  Load the `diamonds.csv` data set and undertake an initial
    exploration of the data. You will find a description of the meanings
    of the variables on the relevant [Kaggle
    page](https://www.kaggle.com/shivam2503/diamonds/)

```{r}
diamonds <- read_csv("4_diamonds_hw/data/diamonds.csv")
```

```{r}
diamonds <- clean_names(diamonds)

head(diamonds)
```
```{r}
summary(diamonds)
```


2.  We expect the `carat` of the diamonds to be strongly correlated with
    the physical dimensions `x`, `y` and `z`. Use `ggpairs()` to
    investigate correlations between these four variables.

```{r}
ggpairs(diamonds, columns = c("carat", "x", "y", "z"))
```



3.  So, we do find significant correlations. Let's drop columns `x`, `y`
    and `z` from the dataset, in preparation to use only `carat` going
    forward.

```{r}
diamonds_trim <- diamonds %>%
  select(-c("x", "y", "z", "x1"))
```

4.  We are interested in developing a regression model for the `price`
    of a diamond in terms of the possible predictor variables in the
    dataset.
    
4.1.  Use `ggpairs()` to investigate correlations between `price` and the
    predictors (this may take a while to run, don't worry, make coffee
    or something).
    
    ```{r}
ggpairs(diamonds_trim)
```


2.  Perform further `ggplot` visualisations of any significant
    correlations you find.

```{r}
alias(lm(price ~ ., data = diamonds_trim))
```

```{r}
diamonds_trim %>% 
ggplot() +
  geom_point((aes(y=carat, x=price, colour = clarity)))
```

```{r}
diamonds_trim %>% 
ggplot() +
  geom_point((aes(y=table, x=price)))
```
```{r}
diamonds_trim %>% 
ggplot() +
  geom_point((aes(y=depth, x=price)))
```

```{r}
diamonds_trim %>% 
ggplot() +
  geom_point((aes(y=depth, x=table)))
```
```{r}
diamonds_trim %>% 
ggplot() +
  geom_point((aes(y=carat, x=price)))
```
5.  Shortly we may try a regression fit using one or more of the
    categorical predictors `cut`, `clarity` and `color`, so let's
    investigate these predictors. Investigate the levels of these
    predictors. How many dummy variables do you expect for each of them?
    
There will be one less dummy variables than there are categories:
    
Cut will produce (5-1) 4 dummy variables
Clarity will produce (8-1) 7 dummy variables
Colour will produce (7-1) 6 dummy variables

```{r}
diamonds_dummy_cut <- diamonds_trim %>%
  fastDummies::dummy_cols(select_columns = "cut", 
                          remove_first_dummy = TRUE, 
                          remove_selected_columns = TRUE) 

head(diamonds_dummy)
```
```{r}
diamonds_dummy_clarity <- diamonds_trim %>%
  fastDummies::dummy_cols(select_columns = "clarity", 
                          remove_first_dummy = TRUE, 
                          remove_selected_columns = TRUE) 

head(diamonds_dummy_clarity)
```
```{r}
diamonds_dummy_color <- diamonds_trim %>%
  fastDummies::dummy_cols(select_columns = "color", 
                          remove_first_dummy = TRUE, 
                          remove_selected_columns = TRUE) 

head(diamonds_dummy_color)
```

```{r}
diamonds_trim %>% 
  ggplot(aes(x = price)) +
  geom_density() +
  geom_vline(xintercept = mean(diamonds_trim$price),
             size = 1,
             colour = "purple") +
  labs(title = "Raw Data")
```

```{r}
price_mean <- mean(diamonds_trim$price)
price_sd <- sd(diamonds_trim$price)

diamonds_stand <- diamonds_trim %>% 
  mutate(price_scaled = (price - price_mean) / price_sd)

diamonds_stand
```



```{r}
diamonds_stand %>% 
  ggplot(aes(x = price_scaled)) +
  geom_density() +
  geom_vline(xintercept = mean(diamonds_stand$price_scaled), size = 1, colour = "red") +
  labs(title = "Standardised data")
```


6.  Start with simple linear regression. Regress `price` on `carat` and
    check the regression diagnostics.

```{r}
# carat is the DEPENDENT variable
# it is described by price
model <- lm(price ~ carat, diamonds_trim)
summary(model)
```

```{r}
diamonds_trim %>% 
  ggplot(aes(x = carat, y = price)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```
```{r}
autoplot(model)
```
```{r}
diamonds_trim %>%
  ggplot(aes(y = price)) + 
  geom_boxplot()
```


7.  Add another predictor of your choice. Check your assumptions,
    diagnostics, and interpret the model.


```{r}
diamonds_trim %>%
  ggplot(aes(x = table, y = price)) + 
  geom_boxplot()
```

```{r}
diamonds_trim %>%
  summarise(cor = cor(table, price))
```

```{r}
model2 <- lm(price ~ carat + table, data = diamonds_trim)
```

```{r}
autoplot(model)
autoplot(model2) 
```

```{r}
summary(model)
summary(model2)
```

------------------------------------------------------------------------

# **2 Extension**

8.  Try adding an interaction between `log(carat)` and your chosen
    categorical predictor. Do you think this interaction term is
    statistically justified?

```{r}
diamonds_trim <- diamonds_trim %>% 
  mutate(log_carat = log(carat))
```

```{r}
# log_carat is the DEPENDENT variable
# it is described by price
model <- lm(price ~ log_carat, diamonds_trim)
summary(model)
```

```{r}
diamonds_trim %>% 
  ggplot(aes(x = log_carat, y = price)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
autoplot(model)
```

9.  Find and plot an appropriate visualisation to show the effect of
    this interaction
