---
title: "Programming in R Homework - Answers"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

```{r}
library(tidyverse)
```


# MVP


# **Question 1.** ----
Find the weather forecast for the next three days online. Create a list of lists in which each inner list should be labelled with the name of the day of the week. Each inner list should contain named elements telling you the temperature, chance of rain and wind speed.

```{r}
# Data for the next 3 days
# Temperature, Chance of Rain, Wind Speed
# Taken for noon for each day (forecast)
weather_3day <- list(
    "Thursday" = list(temperature = 5, chance_of_rain = 13, 
                        wind_speed = 9),
    "Friday" = list(temperature = 4, chance_of_rain = 17, 
                        wind_speed = 11),
    "Saturday" = list(temperature = 5, chance_of_rain = 24, 
                        wind_speed = 12)
)
weather_3day
```

# **Question 2.** ----
Using the list you've just made find:

  i. The chance of rain tomorrow (index the list)
  ii. The wind speed on the day after tomorrow
  iii. The temperature tomorrow.

```{r}
# The chance of rain tomorrow
weather_3day$Thursday$chance_of_rain

# The wind speed on the day after tomorrow
weather_3day$Friday$wind_speed

# The temperature tomorrow.
weather_3day$Thursday$temperature
```

# **Question 3.** ----
Find a 14 day forecast online and create a matrix of the forecasts for temperature, where each week is one row and columns correspond to days.

You will want to give a name to the matrix you create.

```{r}
forecast_14day <- cbind(c(5, 4, 5, 7, 6, 7, 8),c(9, 8, 6, 6, 5, 5, 5))
colnames(forecast_14day) <- c("Week1", "Week2")
rownames(forecast_14day) <- c("Day1", "Day2", "Day3", "Day4", 
                              "Day5", "Day6", "Day7")
forecast_14day
```

# **Question 4.** ----
From the matrix you just made select:

  i. The third day of the second week
  ii. The whole second week
  iii. The first, third and fifth days in both weeks.


```{r}
# The third day of the second week
forecast_14day["Day3", "Week2"]

# The whole second week
forecast_14day[, "Week2"]

# The first, third and fifth days in both weeks
forecast_14day[c("Day1", "Day3", "Day5"), ]

```


# **Question 5.** ----
Write a function that pre-appends "The answer is: " to any character input, while for numeric input, it either just returns the number for numbers less than 10, or divides the number by 10 for numbers greater than 10. For anything else return "Invalid input".



```{r}
# Got really confused with my ifs and elses and brackets
# This is not working and I need to step away!!
# I was pairing with Emily on this so you may see some similar
# code on this question - but not getting much further tonight
x <- 30
if (class(x) == "numeric" & (x) < 10){
      print ("x")
  } if {
    (class(x) == "numeric" & (x) >= 10)
      print (x/10)
    } else {
  if (class(x) == "character"){
    paste("The answer is", "x") 
    else 
      print("Invalid Input")
}
```

```{r}
# Answers from homework
answer_or_divide <- function(x) {
  if (is.character(x)) {
    return(paste("The answer is: ", x))
  }

  if (is.numeric(x)) {
    if (x < 10) {
      return(x)
    } else {
      return(x / 10)
    }
  }

  return("Invalid input")
}
```

Testing:

```{r}
answer_or_divide("forty-two")
answer_or_divide("42")
answer_or_divide(9)
answer_or_divide(19)
```

```{r}
# Notes from homework review
# check if the number is greater or less than 10
# check if the input is a character
# check if the input is a numeric
# using a return() - which is a key word function - it stops the execution of the 
# function and returns the answer

answer_or_divide <- function(x) {
  if (is.character(x)) {
    return(paste("The answer is: ", x))
  }

  if (is.numeric(x)) {
    if (x < 10) {
      return(x)
    } else {
      return(x / 10)
    }
  }

  return("Invalid input")
}
```


# **Question 6.** ----
Write a function which checks if two vectors have the same elements. 

* If they have the same elements in the same order, return "Exactly the same"
* If they have the same elements but in a different order, return "The same"
* If they don't have the same elements, return "Different".

# **Hints** ----
* You can check if two vectors have the same elements but in a different order by using `sort` to sort each vector first and then checking if they are equal.
* Have a look at the `all()` function

```{r}
is_same_vector <- function(vector_1, vector_2) {
  if (all(vector_1 == vector_2)) {
    return("Exactly the same")
  }

  if (all(sort(vector_1) == sort(vector_2))) {
    return("The same")
  }

  return("Different")
}
```

```{r}
a <- c(1, 2, 3, 4, 5)
b <- c(1, 4, 2, 3, 5)
c <- c(9, 4, 8, 1, 0)
d <- c(1, 2, 3, 4, 5)
```

```{r}
is_same_vector(a, a)
is_same_vector(a, b)
is_same_vector(a, c)
is_same_vector(a, d)
```


## Extension ----


# **Question 7.** ----
Write code that prints a multiplication table for numbers up to 12. i.e. there will be 12 rows and 12 columns, and each element should be the row number multiplied by the column number.

*From: https://adriann.github.io/programming_problems.html*


```{r}
m <- matrix(nrow = 12, ncol = 12)

num_rows <- dim(m)[1]
num_cols <- dim(m)[2]

for (row_index in 1:num_rows) {
  for (col_index in 1:num_cols) {
    m[row_index,col_index]=row_index*col_index
  }
  
}
colnames(m) <- c(1:12)
rownames(m) <- c(1:12)
m

```


# **Question 8.**  -----
Write a function that takes a vector, and creates a text art histogram, like the one below.

You may want to use the function `cat()`. This is similar to `print()`, but prints directly, without vector markers. You can also include newlines `"\n"` when using `cat()`.

```
1 | ***
2 | *
3 | *******
4 | ****
5 | **
```

```{r}
make_histogram <- function(x){
  
  distinct <- sort(unique(x))
  
  for (i in distinct){
    
    count <- sum(x == i)
    
    cat(i)
    cat(" | ")
    cat(rep("*", count))
    cat("\n")
    
  }
  
}
```

```{r}
numbers <- sample(1:5, size = 50, replace = TRUE) 
numbers
```

```{r}
make_histogram(numbers)
```

```{r}
cat(paste(letters, 100* 1:26), fill = TRUE, labels = paste0("{", 1:10, "}:"))

```

