---
title: "R Notebook"
output: html_notebook
---

# Question 1 ----

After you’ve read in the data and assigned it to a variable with a useful name, investigate it. Write down some ideas for what you could do to further investigate/clean this data using what you’ve learned today (no need to code it though, just jot down thoughts and ideas). Feel free to also write down any general thoughts you get looking at the data - this task is just to get you thinking about data and data cleaning/prepping.

# Import the data ----
```{r}
library(tidyverse)
movies <-read_csv("movies.csv")
```

# Look at the data

Using glimpse(), ls() and dim()
Note: dim() does not add anything to glimpse()

The data has 4,803 rows and 13 columns
The data is a mix of characters and numbers

The first two columns appear to be the same - would possibly need to do a 
check to see if this is the case for all of the data
    On further examination - the difference appears to be where the film
    is not originally written in English so the original_title gives
    the original name of the film and the title column is the translated
    version.
    
The data does not appear to be sorted
The budget field is formatted in a strange way
THere are some gaps in the data
The overview column is very long and perhaps useful for reference but not 
of any data use.


```{r}
glimpse(movies)
```

```{r}
ls(movies)
```

```{r}
dim(movies)
```

# Question 2 ----

A lot of interesting stuff in there! But for now, we are only interested in the movies’ titles, their runtime as well as their budget. Select the three columns that contain this information and save them into a new variable called movies_selected. Hint: you will have to make a choice between original_title and title.

```{r}
movies_selected <- movies %>% 
  select(title,runtime,budget)
movies_selected
```



# Question 3 ----

Ok, now take your subsetted data movies_selected, and count the number of missing values in each column.

```{r}
movies_selected %>%
  summarise (count=sum(is.na(title)))
```

```{r}
movies_selected %>%
  summarise (count=sum(is.na(budget)))
```

```{r}
movies_selected %>%
  summarise (count=sum(is.na(runtime)))
```

Runtime has two NAs, budget and title have none

# Question 4 ----

There are not too many NAs, but there are actually quite a few movies with a runtime of 0 in this dataset. Since it’s impossible for a movie to have a runtime of 0, use the function na_if to convert runtimes of length 0 into NAs. Save this into a new variable called movies_runtime. Then count how many missing values there are now in the column runtime.

```{r}
movies_selected %>%
  mutate (movies_runtime = na_if(runtime, "0")) %>% 
  arrange(runtime) %>% 
  summarise (count=sum(is.na(runtime)))
```

There are now 2 missing values in the movies_runtime column

# Question 5 ----

Impute the missing values in runtime with the median runtime using coalesce and save this into a new variable called movies_imputed. There should now be no missing values in the column runtime - check this to make sure.

```{r}
movies_imputed <- movies_selected %>%
  mutate(runtime = coalesce(runtime,median(runtime,na.rm = TRUE)))  
 #filter(is.na(runtime))
```

# Question 6 ----

Finally, use the slice_min and slice_max functions on movies_imputed to get the movies with the 10 shortest and 10 longest runtimes.

```{r}

movies_imputed %>% 
  mutate(runtime = coalesce(runtime,median(runtime,na.rm = TRUE)))  %>% 
  slice_max(runtime, n = 10)
```

```{r}
movies_imputed %>% 
  mutate(runtime = coalesce(runtime,median(runtime,na.rm = TRUE)))  %>% 
  slice_min(runtime, n = 10)
```

# Question 7 ----

Using the if_else function, impute anything with a budget below $100 with the median budget. 
Overwrite your movies_imputed so it contains the new budget values.

I know I am going wrong with this bit of code - so going to leave it until we have the 
answers to work out where I am going wrong
