---
title: "R Notebook"
output: html_notebook
---
#Import and View Data

```{r}
library(RXKCD)
library(beepr)
library(tidyverse)
library(CodeClanData)
library(dplyr)
```

# Error 1 ----

Create some_letters which is a character vector containing the following letters: “E”, “R”, “R”, “O” and “R”.

                      ^
```{r}
some_letters <- c("E", "R", "R", "O", "R")
##Error: <text>:1:21: unexpected ','
## 1: some_letters <- ("E",
##   
```

# Error 2 ----

Create message which is a vector containing the following strings: “does”, “this”, “work” and “?”, then convert the vector to uppercase using toupper.

message <- c("does", "this" "work", "?")

                              ^
```{r}
message <- c("does", "this", "work", "?")
toupper(message)
## Error: <text>:1:29: unexpected string constant
## 1: message <- c("does", "this" "work"
##   
```

# Error 3

Select columns mpg, cyl, and disp from mtcars and keep only rows where mpg is less than 20..

```{r}
mtcars %>% 
  select(mpg, cyl, disp) %>%
  filter(mpg < 20)
## Error in mtcars %>% select(mpg, cyl, disp) %>>% filter(mpg < 20): could not find function "%>>%"   
```
                 
# Error 4 ----
Select columns Sepal.Length, Sepal.Width, and Species and arrange by Sepal Length.

```{r}
iris %>% 
  select(Sepal.Length, Sepal.Width, Species) %>% 
  arrange(Sepal.Length)
## Error in iris %>% select(Sepal.Width, Species) %>% arrange(Sepal.Length): could not find function "%>%"
             
```

# Error 5 ----

Find the average tooth length for where the supplement is vitamin C.

```{r}
ToothGrowth %>%
  group_by(supp) %>%
  summarise(avg_len_vc = mean(len)) %>% 
  filter(supp == "VC")
## Error in ToothGrowth %>% summarise(avg_len_vc = mean(len)) %>% filter(supp == : could not find function "%>%"
```

# Error 6 ----
Create a vector of animals (cat, dog, rabbit, horse), then use the sample function to get 10 random animals from those categories.

```{r}
animals <- c("cat", "dog", "rabbit", "horse")
## Error in eval(expr, envir, enclos): object 'dog' not found
sample(animals, 10, replace = TRUE)
## Error in sample(animals, 10): object 'animals' not found
```

# Error 7 ----
Create dog_names and cat_names, then use intersect to check which common pet names are shared by both dogs and cats.

```{r}
dog_names <- c("Lucy", "Molly", "Charlie", "Luna", "Bella", "Cooper")

cat_names <- c("Luna", "Lily", "Lucy", "Charlie", "Max", "Oliver")
## Error in eval(expr, envir, enclos): object 'cat_names' not found
intersect(dog_names, cat_names)
## Error in Intersect(dog_names): could not find function "Intersect"
```



# Error 8 ----
Filter weights equal to or below 150 grams, then create a new column ‘weight_kg’ that contains the weight in kilograms.

                 
```{r}
chickwts %>% 
  filter(weight <= 150) %>% 
  mutate(weight_kg = weight * 0.001)
## Error: <text>:2:18: unexpected '<'
## 1: chickwts %>% 
## 2:   filter(weight =<
##    
```


# Error 9 ----
Create a tibble with column alphabet which contains the English alphabet and column numbers which has a unique number for each letter. Then remove rows where the letter is A, B or C.

```{r}
tibble(alphabet = LETTERS,
       numbers = 1:26) %>% 
  filter(alphabet != c("A", "B", "C"))
## Error in tibble(alphabet = LETTERS(), numbers = 1:25) %>% filter(letter != : could not find function "%>%"
```



# Error 10 ----

The chunk below creates the data for this error (don’t change anything in it):
Don't change anything in this code chunk
but you will need to include it in your notebooks

set.seed(999)

```{r}
scent_ratings <- tibble(
  iD = rep(1:10, each = 5),
  SCENT = rep(c("New book", "Cut grass", "Rose oil", "Fresh linen", "Wet dog"), 10),
  rating = sample(c(0:5, NA), 50, replace = TRUE),
  PersonAge = rep(NA, 50)
)
```

```{r}
## Error in tibble(iD = rep(1:10, each = 5), SCENT = rep(c("New book", "Cut grass", : could not find function "tibble"

##Clean the variable names and drop missing ratings from scent_ratings and calculate an average rating per scent.

scent_ratings_new <- scent_ratings %>% 
  janitor::clean_names()

## Error in scent_ratings %>% janitor:clean_names(): could not find function "%>%"
scent_ratings_new %>% 
  select(i_d, scent, rating) %>% 
  drop_na() %>% 
  group_by(scent) %>% 
  summarise(mean_rating = mean(rating))
## Error in scent_ratings %>% drop_na() %>% group_by(scent) %>% summarise(mean_rating = mean()): could not find function "%>%"
```