---
title: "R Notebook"
output: html_notebook
editor_options: 
  markdown: 
    wrap: 72
---

```{r}
library(tidyverse)
library(modelr)
library(GGally)
library(ggfortify)
```

```{r}
house_data <- read_csv("data/kc_house_data.csv")
```

# Question 1: Exploratory

```{r}
summary(house_data)
```

There are no missing values

Columns to be removed: id date sqft_living15 sqft_lot15 zipcode

Columns to be changed: waterfront yr_renovated - to renovated

Considerations: condition - regrouped to 5 bands grade - left as is

```{r}
house_data <- house_data %>% 
  mutate(renovated = ifelse(yr_renovated > 0, TRUE, FALSE ))
```

```{r}
house_data <- house_data %>% 
  mutate(waterfront = ifelse(waterfront > 0, TRUE, FALSE ))
```

```{r}
house_data <- house_data %>% 
  mutate(graded = case_when(grade >= 11 ~ 'high quality',
                   grade >= 8 ~ 'above average',
                   grade >= 7 ~ 'average',
                   grade >= 4 ~ 'below average',
                   TRUE ~ 'falls short'
          ))
```

```{r}
house_data %>% 
  group_by(condition) %>% 
  summarise(count = n())
```

```{r}
house_data %>% 
  group_by(grade) %>% 
  summarise(count = n())
```
```{r}
# From homework answers
# tidy up data. In particular treat condition and grade as factor, as they are
# ordinal categorical
houses_tidy <- houses %>%
  select(-c("id", "date", "sqft_living15", "sqft_lot15", "zipcode")) %>%
  mutate(waterfront = as.logical(waterfront)) %>%
  mutate(renovated = yr_renovated != 0) %>%
  select(-"yr_renovated") %>%
  mutate(condition = as_factor(condition)) %>%
  mutate(grade = as_factor(grade))

glimpse(houses_tidy)
```

```{r}
# Remove extra columns
house_trim <- house_data %>% 
  select(-c(renovated, id, date, sqft_living15, sqft_lot15, zipcode))
```

# Question 2

Check for aliased variables using the alias() function (this takes in a
formula object and a data set). \[Hint - formula price \~ . says 'price
varying with all predictors', this is a suitable input to alias()\].
Remove variables that lead to an alias. Check the 'Elements of multiple
regression' lesson for a dropdown containing further information on
finding aliased variables in a dataset.

```{r}
alias(lm(price ~ ., data = house_trim))
```

So sqft_basement is the same as sqft_living - sqft_above - so we can
remove two of these

```{r}
# Remove extra columns
house_trim <- house_trim %>% 
  select(-c(sqft_living, sqft_basement))
```

```{r}
# From homework answers
# seems that sqft_basement can be computed from sqft_living - sqft_above.
# let's drop sqft_living leaving just the two contributions sqft_basement and 
# sqft_above
houses_tidy <- houses_tidy %>%
  select(-"sqft_living")

glimpse(houses_tidy)
```


# Question 3

Systematically build a regression model containing up to four main
effects (remember, a main effect is just a single predictor with
coefficient), testing the regression diagnostics as you go * splitting
datasets into numeric and non-numeric columns might help ggpairs() run
in manageable time, although you will need to add either a price or
resid column to the non-numeric dataframe in order to see its
correlations with the non-numeric predictors.

```{r}
house_trim %>% 
GGally::ggpairs()
```
Ooooft - that look a bit crazy

```{r}
houses_tidy_numeric <- house_trim %>%
  select_if(is.numeric)

houses_tidy_nonnumeric <- house_trim %>%
  select_if(function(x) !is.numeric(x))

houses_tidy_nonnumeric$price <- house_trim$price

ggpairs(houses_tidy_numeric)
ggpairs(houses_tidy_nonnumeric)
```

```{r}

cor(houses_tidy_numeric) %>%
  as_data_frame(rownames = "var1") %>%
    pivot_longer(cols = -var1, names_to = "var2", values_to = "coeff") %>%
    filter(var1 != var2) %>%
    filter(var1 == "price") %>% 
    arrange(desc(abs(coeff)))
```

Looking at all of the different combinations I think the variables I want to consider are:
grade
sqft_above
bathrooms
view
waterfront

## Building the model - first effect (comparing waterfront and grade)
First I am going to try waterfront (the best fitting non-numeric variable) and grade (the best fitting numeric variable) and see how they compare

```{r}
mod1a <- lm(price ~ waterfront, data = house_trim)
mod1a
```

```{r}
mod1b <- lm(price ~ grade, data = house_trim)
mod1b
```

Woah - so waterfront has a *huge* effect on price.

For every $1 increase in price waterfront goes up 1.1 million with an intercept of $531564

For grade the intercept starts much lower at -$1056045 and then goes up $208458

So - let's start with waterfront

```{r}
summary(mod1a)
```

Hmm - the R-Squared value is low at only 7% of prices being affected by waterfront
So perhaps it is a fantastic indication of price but is only found in a few properties.

Lets have a quick look

```{r}
house_trim %>% 
  group_by(waterfront) %>% 
  summarise(count = n())
```

Only 163 of the properties are waterfront (0.75%) - so it is important for price - but overall has little impact on the rest of the sales

```{r}
par(mfrow = c(2,2))
plot(mod1a)
```

We should definitely keep it in as a model predictor

Now - have a better look at grade:

```{r}
mod1b <- lm(price ~ grade, data = house_trim)
mod1b
```

```{r}
summary(mod1b)
```

The Rsquared value is much higher: 44.55% of the prices are affected by grade and with a very small p-value this is also significant

```{r}
par(mfrow = c(2,2))
plot(mod1b)
```

```{r}
autoplot(mod1b)
```

## Adding in the second variable

```{r}
mod1a <- lm(price ~ waterfront, data = house_trim)
mod1a
```

```{r}
house_remaining_resid <- house_trim %>%
  add_residuals(mod1a) %>%
  select(-c("price", "waterfront"))
house_remaining_resid
```

```{r}
house_remaining_resid %>%
GGally::ggpairs()
```

```{r}
mod1d <- lm(resid ~ grade, data = house_remaining_resid)
mod1d
```

Resid is our predicted value

Resid = -1543405 + (201571 x grade)

For every point increase in resid grade goes up by 201571

```{r}
#Re -running with correct column
mod1da <- lm(resid ~ graded, data = house_remaining_resid)
mod1da
```


```{r}
# Comparing two versions of grade/graded
summary(mod1d)
summary(mod1da)
```

44.8% of property prices are affected by waterfront and grade
37.9% if it is split into categories - so keeping it as individual numbers is more efficient

```{r}
par(mfrow = c(2,2))
plot(mod1d)
```

## Adding the next one in

```{r}
mod2b <- lm(price ~ waterfront + grade + sqft_above, data = house_trim)
summary(mod2b)
```

51% can now be predicted with waterfront, grade and sqft

```{r}
anova(mod2b, mod1d)
```

## Adding the last one 

```{r}
mod2c <- lm(price ~ waterfront + grade + sqft_above + bathrooms, data = house_trim)
summary(mod2c)
```

```{r}
anova(mod2c, mod2b)
```

Final linear model:
```{r}
lm(formula = price ~ waterfront + grade + sqft_above + bathrooms, 
    data = house_trim)
```


51.6%


