---
title: "Text Mining Homework"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    css: ../../../styles.css
  pdf_document: default
---

# MVP

Using the dataset `austen_books()` from the package `janeaustenr`:

1.  Find the most common words in both Pride & Prejudice and Sense & Sensibility.
2.  Find the most common words in both Pride & Prejudice and Sense & Sensibility, not including stop words.
3.  Find the most common sentiment words in both Pride & Prejudice and Sense & Sensibility.

# Extension

Taking your results above. Can you create a plot which visualises the differences between the books?

```{r}
library(tidyverse)
library(tidytext) # Julia Silge
library(janeaustenr)
library(textdata)
library(ggwordcloud)
library(wordcloud)
library(wordcloud2)
library(RColorBrewer)
library(janitor)
```

```{r}
# Import the data set for all books
all_books <- austen_books()
```

```{r}
# Check the names of the books in the main table
all_books %>% 
  group_by(book) %>% 
  summarise(count = n())
```

# 1.  Find the most common words in both Pride & Prejudice and Sense & Sensibility.

```{r}
# Top words in Pride and Prejudice
pride_top_all <- all_books %>%
  filter(book == "Pride & Prejudice") %>%
  unnest_tokens(word, text) %>%
  count(word, sort = TRUE) 
pride_top_all
```

```{r}
# Top words in Sense and Sensibility
sense_top_all <- all_books %>%
  filter(book == "Sense & Sensibility") %>%
  unnest_tokens(word, text) %>%
  count(word, sort = TRUE) 
sense_top_all
```

# 2.  Find the most common words in both Pride & Prejudice and Sense & Sensibility, not including stop words.

Filter the required books and remove stop words

```{r}
# Top words in Pride and Prejudice
pride <- all_books %>%
  filter(book == "Pride & Prejudice") %>%
  unnest_tokens(word, text) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words)
pride
```

```{r}
# Top words in Sense and Sensibility
sense <- all_books %>%
  filter(book == "Sense & Sensibility") %>%
  unnest_tokens(word, text) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words)
sense
```

```{r}
wordcloud2(data=pride)
```

```{r}
wordcloud2(data=sense)
```

# 3.  Find the most common sentiment words in both Pride & Prejudice and Sense & Sensibility.

```{r}
# For some reason this code which I found in github creates the tables in a useable format
all_books2 <- austen_books() %>%     
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", 
                                                 ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```

```{r}
# Import sentiment libraries
afinn <- get_sentiments("afinn")
bing <- get_sentiments("bing")
loughran <- get_sentiments("loughran")
nrc <- get_sentiments("nrc")
```

```{r}
# Remind myself of the sentiments in each one
bing %>% 
  group_by(sentiment) %>% 
  summarise(count = n())

loughran %>% 
  group_by(sentiment) %>% 
  summarise(count = n())

nrc %>% 
  group_by(sentiment) %>% 
  summarise(count = n())
```

All books - positive and negative

```{r}
all_books2 %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(random.order = FALSE, colors = c("dark blue", "firebrick1"),
                   max.words = 120)
```

```{r}
all_books2 %>%
  filter(book == "Pride & Prejudice") %>% 
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(random.order = FALSE, colors = c("dark blue", "light blue"),
                   max.words = 120)

all_books2 %>%
  filter(book == "Sense & Sensibility") %>% 
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(random.order = FALSE, colors = c("blueviolet", "cyan"),
                   max.words = 120)

all_books2 %>%
  filter(book == "Mansfield Park") %>% 
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(random.order = FALSE, colors = c("darkorchid", "deeppink"),
                   max.words = 120)

all_books2 %>%
  filter(book == "Persuasion") %>% 
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(random.order = FALSE, colors = c("deepskyblue4", "deepskyblue"),
                   max.words = 120)

all_books2 %>%
  filter(book == "Northanger Abbey") %>% 
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(random.order = FALSE, colors = c("chocolate1", "coral1"),
                   max.words = 120)

all_books2 %>%
  filter(book == "Emma") %>% 
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(random.order = FALSE, colors = c("forestgreen", "darkolivegreen1"),
                   max.words = 120)
```



I am going to use the nrc database for individual books:


```{r}
pride_sentiments <- all_books2 %>% 
  filter(book == "Pride & Prejudice") %>% 
  inner_join(nrc)

pride_sentiments %>% 
  filter(sentiment == "positive") %>% 
  count(word, sort = TRUE)
```




```{r}
# Positive codes from sense and sensibility
sense_sentiments <- all_books2 %>% 
  filter(book == "Sense & Sensibility") %>% 
  inner_join(nrc)

sense_sentiments %>% 
  filter(sentiment == "positive") %>% 
  count(word, sort = TRUE)
```

```{r}
pride_sentiment_summary <- pride_sentiments %>% 
# filter(sentiment == "positive") %>% 
  group_by(sentiment) %>% 
  count(word, sort = TRUE)

sense_sentiment_summary <- sense_sentiments %>% 
# filter(sentiment == "positive") %>% 
  group_by(sentiment) %>% 
  count(word, sort = TRUE)
```

```{r}
# Top 5 sentiments for each of the 10 sentiment values - P&P
pride_top_sentiments <- pride_sentiment_summary %>%
  arrange_(~ desc(n)) %>%
  group_by_(~ sentiment) %>%
  slice(1:5)
pride_top_sentiments
```

```{r}
# Top 5 sentiments for each of the 10 sentiment values - S&S
sense_top_sentiments <- sense_sentiment_summary %>%
  arrange_(~ desc(n)) %>%
  group_by_(~ sentiment) %>%
  slice(1:5)
sense_top_sentiments
```

# Story Arcs

```{r}
afinn_pride_sentiments <- pride %>%
  inner_join(afinn)
```

```{r}
# Looking at the story arcs
# Not sure what is going on with Persuasion though!
all_books %>%
  filter(book %in% c("Sense & Sensibility", "Pride & Prejudice", "Mansfield Park", 
                     "Persuasion", "Northanger Abbey", "Emma")) %>%
  unnest_tokens(word, text) %>%
  inner_join(get_sentiments("afinn")) %>%
  group_by(book) %>%
  mutate(
    word_n = row_number()
  ) %>%
  mutate(
    story_position = word_n/max(word_n) 
  ) %>% 
ggplot +
  aes(x = story_position, y = value, colour = book) +
  geom_smooth(se = FALSE) +
  guides(colour = FALSE) +
  facet_wrap(~ book, nrow = 3)
```


```{r}
# Looking at the story arcs
# Not sure what is going on with Persuasion though!
all_books2 %>%
  filter(book %in% c("Sense & Sensibility", "Pride & Prejudice", "Mansfield Park", 
                     "Persuasion", "Northanger Abbey", "Emma")) %>%
  #unnest_tokens(word, text) %>%
  inner_join(get_sentiments("afinn")) %>%
  group_by(book) %>%
  mutate(
    word_n = row_number()
  ) %>%
  mutate(
    story_position = word_n/max(word_n) 
  ) %>% 
ggplot +
  aes(x = story_position, y = value, colour = book) +
  geom_smooth(se = FALSE) +
  guides(colour = FALSE) +
  facet_wrap(~ book, nrow = 3)
```

```{r}
# Looking at the story arcs
# Not sure what is going on with Persuasion though!
all_books2 %>%
  filter(book %in% c("Persuasion")) %>%
  #unnest_tokens(word, text) %>%
  inner_join(get_sentiments("afinn")) %>%
  group_by(book) %>%
  mutate(
    word_n = row_number()
  ) %>%
  mutate(
    story_position = word_n/max(word_n) 
  ) %>% 
ggplot +
  aes(x = story_position, y = value, colour = book) +
  geom_smooth(se = FALSE) +
  guides(colour = FALSE) 
 # facet_wrap(~ book, nrow = 3)
```
