---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
library(broom)
library(modelr)
library(caret)
library(leaps)
library(GGally)
library(ggfortify)
```

```{r}
raw_avocado <- read_csv("data/avocado.csv")
```

# MVP

We've looked at a few different ways in which we can build models this week, including how to prepare them properly. This weekend we'll build a multiple linear regression model on a dataset which will need some preparation. The data can be found in the data folder, along with a data dictionary

We want to investigate the avocado dataset, and, in particular, to model the AveragePrice of the avocados. Use the tools we've worked with this week in order to prepare your dataset and find appropriate predictors. Once you've built your model use the validation techniques discussed on Wednesday to evaluate it. Feel free to focus either on building an explanatory or a predictive model, or both if you are feeling energetic!

As part of the MVP we want you not to just run the code but also have a go at interpreting the results and write your thinking in comments in your script.

Hints and tips

region may lead to many dummy variables. Think carefully about whether to include this variable or not (there is no one 'right' answer to this!) Think about whether each variable is categorical or numerical. If categorical, make sure that the variable is represented as a factor. We will not treat this data as a time series, so Date will not be needed in your models, but can you extract any useful features out of Date before you discard it? If you want to build a predictive model, consider using either leaps or glmulti to help with this.

# Exploratory Data Analysis

```{r}
summary(raw_avocado)
```

We have 18248 rows and 14 variables

1.  x1 - Row count - this can be removed
2.  Date - We will not treat this data as a time series, so Date will not be needed in your models, but can you extract any useful features out of Date before you discard it? (How about month?) runs from 2015-2018
3.  Ave Price - this is the value we will be modelling/predicting - ave price of a single avocado
4.  Total Volume - total number of avocadoes
5.  4046: Small/Medium Hass Avocado
6.  4225: Large Hass Avocado
7.  4770: Extra Large Hass Avocado
8.  Total Bags
9.  Small Bags
10. Large Bags
11. XLarge Bags
12. type: conventional or organic
13. year: the year
14. region: the city or region of the observation

```{r}
# Clean Names
raw_avocado <- raw_avocado %>% 
clean_names()
```

```{r}
# Fix the date field as it is not currently a date field
raw_avocado<- raw_avocado %>%
  mutate(date= ymd(date))
```

```{r}
# Add in a month column
raw_avocado<- raw_avocado %>%
  mutate(month = month(date, label = TRUE, abbr = FALSE))
```

```{r}
raw_avocado %>% 
  group_by(month) %>% 
  summarise(count=n())
```

Perhaps group the Months into quarters

```{r}
# Add in a quarter column
raw_avocado<- raw_avocado %>%
  mutate(quarter = quarter(date))
```

```{r}
# Box plot comparing type (conventional vs organic)
ggplot(raw_avocado, aes(x=as.factor(type), y=average_price)) + 
    geom_boxplot(fill="slateblue", alpha=0.2) + 
    xlab("cyl")
```

So the organic avocadoes drive the price up

```{r}
# Simple line graphs looking at some of the variables
ggplot(raw_avocado, aes(x=average_price)) + 
  geom_line(aes(y = x4225), color = "orange", alpha = 0.4) +
  geom_line(aes(y = x4046), color = "darkred", alpha = 0.4) +
  geom_line(aes(y = x4770), color="steelblue", alpha = 0.4) 

```

Doesn't really tell us much - but we get an idea of the shape of the data.

```{r}
regions <- raw_avocado %>% 
  group_by(region) %>% 
  summarise(count = n())
```

There are 54 regions, with the same number of observations from each. For modelling this could be a problem - but perhaps we can find one or two regions that are key for driving up prices.

Perhaps we should look at some simple stats per region.

```{r}
regions <- raw_avocado %>% 
  group_by(region) %>% 
  summarise(count = n(), mean(average_price), mean(x4046), mean(x4225), 
            mean(x4770))
regions
```

```{r}
raw_avocado %>%
  ggplot(aes(x = average_price, y = region)) +
  geom_boxplot()
```
Phew - what a mess

Let's rotate it

```{r}
raw_avocado %>%
  ggplot(aes(x = region, y = average_price)) +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45))
```
Ugly graph - but gives us a glimpse at the variation between regions - so perhaps this is important after all.

```{r}
# Tidy up variables
# Remove row count, date and month
avocado_trim <- raw_avocado %>% 
  select(-c(x1, date, month))
```

# Start Modelling
## Check for aliased variable
```{r}
alias(lm(average_price ~ ., data = avocado_trim))
```

Looks like we have no aliased variables - we are good to go

## Run ggpairs
```{r}
# This causes errors because of the regions
avocado_trim %>% 
GGally::ggpairs()
```


```{r}
# Let's see if it works if we convert to numeric/non-numeric
avocado_trim_numeric <- avocado_trim %>%
  select_if(is.numeric)

avocado_trim_nonnumeric <- avocado_trim %>%
  select_if(function(x) !is.numeric(x))

avocado_trim_nonnumeric$price <- avocado_trim$price

ggpairs(avocado_trim_numeric)
ggpairs(avocado_trim_nonnumeric)
```
So - some observations: Regions continue to cause problems - so need to rethink it. The quarters are being recognised as numeric, not categories - so need to recode

## Recode problem data

```{r}
avocado_trim <- avocado_trim %>% 
  mutate(quarter = str_c("Q", quarter))
```


```{r}
# Remove regions
avocado_trim_nr <- avocado_trim %>% 
  select(-c(region))
```

```{r}
# Attempt two
avocado_trim_numeric <- avocado_trim_nr %>%
  select_if(is.numeric)

avocado_trim_nonnumeric <- avocado_trim_nr %>%
  select_if(function(x) !is.numeric(x))

avocado_trim_nonnumeric$average_price <- avocado_trim_nr$average_price

ggpairs(avocado_trim_numeric)
ggpairs(avocado_trim_nonnumeric)
```
Non-numeric
  Type is definitely a key variable
  Quarter has some influence
  
Numeric correlations (of average price)
Year 0.093
xl bags -0.118
large bags -0.173
x4225 -0.173
small bags -0.175
total bags -0.177
x4770 -0.179
total volume -0.193
x4046 -0.208

The highest correlation scores (top three)
x4046 -0.208
total volume -0.193
x4770 -0.179

## Try exhaustive modelling
to identify key variables

```{r}
# exhaustive selection
regsubsets_exhaustive <- regsubsets(average_price ~ ., 
                                 data = avocado_trim_nr, 
                                 nvmax =8, # maxm size of subsets
                                 method = "exhaustive")
```

```{r}
sum_regsubsets_exhaustive <- summary(regsubsets_exhaustive)
sum_regsubsets_exhaustive
```

```{r}
sum_regsubsets_exhaustive$which
```

```{r}
plot(regsubsets_exhaustive, scale = "adjr2")
```
```{r}
plot(regsubsets_exhaustive, scale = "bic")
```
```{r}
plot(sum_regsubsets_exhaustive$rsq, type = "b")
```
Interestingly there is no elbow in the plot so there is no clear point at which to stop modelling.

```{r}
plot(sum_regsubsets_exhaustive$bic, type = "b")
```
```{r}
summary(regsubsets_exhaustive)$which[6,]
```

Exhausting modelling suggests to us that the key variables (in order) are:
type (organic)
quarter (03)
quarter(04)
year
quarter(02)

## First Variable selection
### Model 1a - type

```{r}
# model 1a - using type as the variable
model1a <- lm(average_price ~ type, data = avocado_trim_nr)
model1a
```

```{r}
# model 1a - using type as the variable
model1a <- lm(average_price ~ type, data = avocado_trim)
model1a
```

Average Price is our predicted value

Average price = 1.158 + (0.496 x Organic(type))

If an avocado is organic the price of it will increase by 0.496 assuming all other variables remain constant.

```{r}
summary(model1a)
```

The p-value is less than 0.05 so we know this is statistically significant. The R^2 value tells us that 37.9% of the variation in the average price can be accounted by the avocado being organic.

```{r}
par(mfrow = c(2,2))
plot(model1a)
```

Before we accept this as our first variable let's check with our second predictor - quarter 3

### Model 1b - quarter

```{r}
# model 1b - using quarter as the variable
model1b <- lm(average_price ~ quarter, data = avocado_trim_nr)
model1b
```

Average Price is our predicted value

Average price = 1.30660 + (0.20631 x Organic(type))

If an avocado is organic the price of it will increase by 0.496 assuming all other variables remain constant.

```{r}
summary(model1b)
```

The p-value is less than 0.05 so we know this is statistically significant. The R^2 value tells us that 4% of the variation in the average price can be accounted by the avocado being organic.

```{r}
par(mfrow = c(2,2))
plot(model1b)
```

Model1a is definitely a better model than Model1b - so let's choose type for the first variable.

## Second Variable selection

Now we need to rerun the analysis to determine the next variable

```{r}
avocado_rem_resid <- avocado_trim_nr %>%
  add_residuals(model1a) %>%
  select(-c("average_price", "type"))
ggpairs(avocado_rem_resid)
```

Coefficients (of the Resid)
total volume -0.063
x4046 -0.088
x4225 -0.038
x4770 -0.064
total bags -0.055
small bags -0.049
large bags -0.069
xl bags -0.012
year 0.118
quarter - some variation - Do I need to make this a dummy?

```{r}
# exhaustive selection
regsubsets_exhaustive2 <- regsubsets(resid ~ ., 
                                 data = avocado_rem_resid, 
                                 nvmax =8, # maxm size of subsets
                                 method = "exhaustive")
```

```{r}
sum_regsubsets_exhaustive2 <- summary(regsubsets_exhaustive2)
sum_regsubsets_exhaustive2
```

Top variables are  Q3, Q4, Year, Q2

I tried to put region back in but it is still running errors - I may test it anyway

So - let's compare quarter, year and region and see which works best

### Model 2a - quarter

```{r}
# model 2a - using quarter as the variable
# bringing back in the original dataset with regions
model2a <- lm(average_price ~ type + quarter, data = avocado_trim)
model2a
```

Average Price is our predicted value

For Quarter 3 
Average price = 1.05863 + (0.496 x Organic(type) + (0.20631 x Quarter3))

If an avocado is organic and picked in quarter 3 the price of it will increase by 0.496 + 0.20631 assuming all other variables remain constant.

```{r}
summary(model2a)
```

The p-value is less than 0.05 so we know this is statistically significant. The R^2 value tells us that 41.93% of the variation in the average price can be accounted by the avocado being organic.

```{r}
par(mfrow = c(2,2))
plot(model2a)
```
I am liking the Q-Q here 


### Model 2b - year

```{r}
# model 2b - using year as the variable
# bringing back in the original dataset with regions
model2b <- lm(average_price ~ type + year, data = avocado_trim)
model2b
```

Going to stop looking at year now - as it is treating it as a numeric 

### Model 2c - region

```{r}
# model 2c - using quarter as the variable
# bringing back in the original dataset with regions
model2c <- lm(average_price ~ type + region, data = avocado_trim)
model2c
```
Oh wow!! This will take some analysis - so let's look at the summary


```{r}
summary(model2c)
```

The p-value is mostly less than 0.05 but there are some regions where is is greater than 0.05 which could make the data misleading

The R^2 value tells us that 54.73% of the variation in the average price can be accounted by the avocado being organic and by the region it is in

```{r}
par(mfrow = c(2,2))
plot(model2c)
```
### Compare Model 1a, 2a and 2c
Time to use anova to compare the models:

```{r}
anova(model1a, model2a)
```

The null hypothesis here is that the models explain the same amount of response variance. The alternative is that they don't. In this case, we find a p-value less than 0.05, and so we reject the null hypothesis and say that the model including type is significantly better than the model excluding it!

However, the model including region is still better overall (with higher r2), and so we choose region over quarter in this case. But perhaps we can include it as a third variable?

```{r}
anova(model1a, model2c)
```

## Third Variable

```{r}
avocado_rem_resid2 <- avocado_trim %>%
  add_residuals(model2c) %>%
  select(-c("average_price", "type", "region"))
ggpairs(avocado_rem_resid2)
```

Coefficients (of the Resid)
total volume -0.017
x4046 -0.018
x4225 -0.023
x4770 -0.024
total bags -0.005
small bags -0.005
large bags -0.008
xl bags -0.031
year 0.139 - disregard this
quarter - some variation - Do I need to make this a dummy?
