---
title: "R Notebook"
output: html_notebook
---

```{r}
library(prob)
library(gtools)
library(tidyverse)
library(janitor)
```

```{r}
titanic <- read_csv("data/titanic.csv")
```

```{r}

class_count <- titanic %>% 
  group_by(class) %>% 
  summarise(n())

class_count
```

```{r}
titanic <- titanic %>% 
  mutate(
    crew = case_when(
      str_detect(class, "crew") ~ TRUE,
      str_detect(class, "staff") ~ TRUE,
      TRUE ~ FALSE
    ),
    .after = age
  )
```

```{r}
# Produces a 2 x 2 table
class_table <- titanic %>% 
  tabyl(class, survived)

class_table
```

```{r}

class_table <- class_table %>% 
  adorn_totals(where = c("row", "col"))

class_table
```

```{r}
class_table_percent <- class_table %>% 
  adorn_percentages() %>% 
  adorn_title()

class_table_percent
```

## Question

Calculate the conditional probabilities prob(survived∣crew) and prob(survived∣passenger) to settle this dispute.

Have a look at the distinct values of the class variable, this will tell you whether each individual was a passenger (1st, 2nd or 3rd class) or a member of the crew (treat restaurant staff as crew)

## Further Analysis

prob(survived|Crew) -> probability of surviving given that a person is part of the crew prob(survived∣passenger)-> probability of surviving given that a person is a passenger

Calculate the conditional probabilities of:
  prob(survived|1st class | passenger)
  prob(survived|2nd class | passenger)
  prob(survived|3rd class | passenger)
  
```{r}
# probability that person is a 1st class passenger 
first_class <- titanic %>% 
  summarise(prob = sum(class == "1st") / n())
first_class
```

```{r}
# probability that person is a 2nd class passenger 
second_class <- titanic %>% 
  summarise(prob = sum(class == "2nd") / n())
second_class
```

```{r}
# probability that person is a 3rd class passenger 
third_class <- titanic %>% 
  summarise(prob = sum(class == "3rd") / n())
third_class
```

```{r}
# probability that person is a passenger
prob_passenger <- titanic %>% 
  summarise(prob = sum(crew == FALSE) / n())
prob_passenger
```

```{r}
# probability that person survived
prob_survived <- titanic %>% 
  summarise(prob = sum(survived == "yes") / n())
prob_survived
```

```{r}
# probability 1st class and survived
prob_survived_1st_class <- titanic %>% 
    filter(survived == "yes", class == "1st") %>% 
    summarise(prob = n()/ nrow(titanic))
prob_survived_1st_class
```

```{r}
# probability 2nd class and survived
prob_survived_2nd_class <- titanic %>% 
    filter(survived == "yes", class == "2nd") %>% 
    summarise(prob = n()/ nrow(titanic))
prob_survived_2nd_class
```

```{r}
# probability 3rd class and survived
prob_survived_3rd_class <- titanic %>% 
    filter(survived == "yes", class == "3rd") %>% 
    summarise(prob = n()/ nrow(titanic))
prob_survived_3rd_class
```


```{r}
# probability passenger and survived
prob_survived_passenger <- titanic %>% 
    filter(survived == "yes", crew == FALSE) %>% 
    summarise(prob = n()/ nrow(titanic))
prob_survived_passenger
```

```{r}
prob_survived_given_passenger <- 
      prob_survived_passenger / prob_passenger

prob_survived_given_passenger

```

```{r}
prob_survived_given_1st_class <- 
      prob_survived_1st_class / first_class

prob_survived_given_1st_class

```

```{r}
prob_survived_given_2nd_class <- 
      prob_survived_2nd_class / second_class

prob_survived_given_2nd_class

```

```{r}
prob_survived_given_3rd_class <- 
      prob_survived_3rd_class / third_class

prob_survived_given_3rd_class

```

```{r}
# Crew probabilities
deck_crew <- titanic %>% 
  summarise(prob = sum(class == "deck crew") / n())
deck_crew

engineering_crew <- titanic %>% 
  summarise(prob = sum(class == "engineering crew") / n())
engineering_crew

restaurant_staff <- titanic %>% 
  summarise(prob = sum(class == "restaurant staff") / n())
restaurant_staff

victualling_crew <- titanic %>% 
  summarise(prob = sum(class == "victualling crew") / n())
victualling_crew

```

```{r}
# probability crew and survived
prob_survived_deck_crew <- titanic %>% 
    filter(survived == "yes", class == "deck crew") %>% 
    summarise(prob = n()/ nrow(titanic))
prob_survived_deck_crew

prob_survived_engineering_crew <- titanic %>% 
    filter(survived == "yes", class == "engineering crew") %>% 
    summarise(prob = n()/ nrow(titanic))
prob_survived_engineering_crew

prob_survived_restaurant_staff <- titanic %>% 
    filter(survived == "yes", class == "restaurant staff") %>% 
    summarise(prob = n()/ nrow(titanic))
prob_survived_restaurant_staff

prob_survived_victualling_crew <- titanic %>% 
    filter(survived == "yes", class == "victualling crew") %>% 
    summarise(prob = n()/ nrow(titanic))
prob_survived_victualling_crew
```

```{r}
prob_survived_given_deck_crew <- 
      prob_survived_deck_crew / deck_crew

prob_survived_given_engineering_crew <- 
      prob_survived_engineering_crew / engineering_crew

prob_survived_given_restaurant_staff <- 
      prob_survived_restaurant_staff / restaurant_staff

prob_survived_given_victualling_crew <- 
      prob_survived_victualling_crew / victualling_crew

prob_survived_given_deck_crew
prob_survived_given_engineering_crew
prob_survived_given_restaurant_staff
prob_survived_given_victualling_crew
```



The probability of surviving given that a person is part of the crew = 0.2370787

The probability of surviving given that a person is a passenger = 0.3796507

The probability of surviving given that a person is a 1st class passenger = 0.6203704	
The probability of surviving given that a person is a 2nd class passenger = 0.415493
The probability of surviving given that a person is a 3rd class passenger = 0.2552891

The probability of surviving given that a person is deck crew = 0.6515152	
The probability of surviving given that a person is engineering crew = 0.2191358
The probability of surviving given that a person is restaurant staff = 0.04347826	
The probability of surviving given that a person is victualling crew = 0.2180974	

```{r}
class <- c("all passengers", "1st class", "2nd class", "3rd class", "all crew", 
           "deck crew", "engineering crew", "restaurant staff", "victualling crew")
prob <- c(0.3796507, 0.6203704, 0.415493, 0.2552891, 0.2370787, 0.6515152, 0.2191358, 0.04347826, 0.2180974)
group <- c("passengers", "passengers", "passengers", "passengers", 
           "crew", "crew", "crew", "crew", "crew")

prob_survival_by_class <- data.frame(class, prob, group)
```

```{r}
ggplot(prob_survival_by_class, aes(fill=class, y=prob, x=group)) + 
    geom_bar(position="dodge", stat="identity")
```

