---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

#Titanic Homework ## Background

A data set detailing the survival or death of passengers and crew members on the RMS Titanic is given in the data folder for this homework - use this data and not the built-in Titanic dataset in R.

Some time after the disaster a journalist wrote that crew members were more likely to have survived the sinking than passengers. This was disputed by a descendant of one of the surviving crew members, who suggested that the opposite was true, i.e. passengers were more likely to have survived than crew.

## Loading packages and importing data

You may want to use prob and gtools later in the homework, and the tidyverse for this section, but if you load prob after tidyverse, dplyr::filter() will be masked. So load your packages in this order:

```{r}
library(prob)
library(gtools)
library(tidyverse)
library(janitor)
```

```{r}
titanic <- read_csv("data/titanic.csv")
```

## Manipulating Data

```{r}

class_count <- titanic %>% 
  group_by(class) %>% 
  summarise(n())

class_count
```

```{r}
titanic <- titanic %>% 
  mutate(
    crew = case_when(
      str_detect(class, "crew") ~ TRUE,
      str_detect(class, "staff") ~ TRUE,
      TRUE ~ FALSE
    ),
    .after = age
  )
```

```{r}
survived_people <- titanic %>% 
  select("crew", "survived")
```

```{r}
# Produces a 2 x 2 table
table <- survived_people %>% 
  tabyl(crew, survived)

table
```

```{r}

table <- table %>% 
  adorn_totals(where = c("row", "col"))

table
```

```{r}
table <- table %>% 
  adorn_percentages() %>% 
  adorn_title()

table
```

## Question

Calculate the conditional probabilities prob(survived∣crew) and prob(survived∣passenger) to settle this dispute.

Have a look at the distinct values of the class variable, this will tell you whether each individual was a passenger (1st, 2nd or 3rd class) or a member of the crew (treat restaurant staff as crew)

## Hints

You may want to create a crew column containing values TRUE or FALSE to help with this For prob(survived∣crew) think 'what proportion of the crew survived?', and similar for prob(survived∣passenger)

## Analysis

prob(survived\|Crew) -> probability of surviving given that a person is part of the crew prob(survived∣passenger)-> probability of surviving given that a person is a passenger

```{r}
# probability that person is a crew member
prob_crew <- survived_people %>% 
  summarise(prob = sum(crew == TRUE) / n())
prob_crew
```

```{r}
# probability that person is a passenger
prob_passenger <- survived_people %>% 
  summarise(prob = sum(crew == FALSE) / n())
prob_passenger
```

```{r}
# probability that person survived
prob_survived <- survived_people %>% 
  summarise(prob = sum(survived == "yes") / n())
prob_survived
```

```{r}
# probability crew and survived
prob_survived_crew <- survived_people %>% 
    filter(survived == "yes", crew == TRUE) %>% 
    summarise(prob = n()/ nrow(survived_people))
prob_survived_crew
```

```{r}
# probability passenger and survived
prob_survived_passenger <- survived_people %>% 
    filter(survived == "yes", crew == FALSE) %>% 
    summarise(prob = n()/ nrow(survived_people))
prob_survived_passenger
```

```{r}
prob_survived_given_crew <- 
      prob_survived_crew / prob_crew

prob_survived_given_crew

```

```{r}
prob_survived_given_passenger <- 
      prob_survived_passenger / prob_passenger

prob_survived_given_passenger

```

## Summary

The probability of surviving given that a person is part of the crew = 0.2370787

The probability of surviving given that a person is a passenger = 0.3796507

Going back to the original question: Some time after the disaster a journalist wrote that crew members were more likely to have survived the sinking than passengers. This was disputed by a descendant of one of the surviving crew members, who suggested that the opposite was true, i.e. passengers were more likely to have survived than crew.

From the analysis of the survival rates it looks like you were mmuch more likely to survive if you were a passenger, rather than a member of the crew.

# Probability Consolidation Questions

Today was pretty full-on! Here are some problems to help you consolidate what you learned. Feel free to work on paper or in R, whichever you are most comfortable with:

MVP - try four of the problems Extension - try more than four!

## Question 1.

If I roll a six-sided dice once, what is the probability of getting a number less than 5?

```{r}
# Probability of rolling one die
s_one_dice <- rolldie(1) 
```

```{r}
dice_less_than_5 <- s_one_dice %>% 
  filter(X1 < 5)
```

```{r}
prob_less_than_5 <- nrow(dice_less_than_5) / nrow(s_one_dice) 
prob_less_than_5
```

Answer is 0.6666667 or 4/6

## Question 2.

How many outcomes are there in the sample space of an experiment in which we toss three dice?

```{r}
# Outcomes of rolling three dice
s_three_dice <- rolldie(3) 

```

There are 216 possible outcomes (6 x 6 x 6)

## Question 3.

In an experiment where we roll four dice:

how many outcomes are in the event where the total of the four dice equals 5, 6 or 7)? what is the probability of this event? what assumptions do we apply in answering part ii? Hint For part i, rowSums(.) may help you setup up the necessary random variable total (or you can use pivot_longer())

```{r}
# Outcomes of rolling four dice
s_four_dice <- rolldie(4) 
```

```{r}
total_5_6_7 <- s_four_dice %>% 
  filter((X1 + X2 + X3 + X4) %in% c(5, 6, 7))
```

```{r}
prob_5_6_7 <- nrow(total_5_6_7) / nrow(s_four_dice)
prob_5_6_7
```

The probability of the total of the four dice roll = 5, 6 or 7 is 0.02623457 (or 34/1296) We assume that each roll is independent of the next.

## Question 4.

Five chairs are placed in a row in a room, along with five people of differing height. The people sit down randomly on the chairs. What is the probability that they sit in order of decreasing height, left to right?

Hint This is an arrangement where order matters. Of all the possible arrangements, how many are correct?

The total number of possibilities for the people sitting down is 5! which is (5 x 4 x 3 x 2 x 1) = 120 There is only one possible combination of the people sitting in the correct order, by height. So - the possibility of this happening is 1/120 which is 0.00833333

## Question 5.

I roll six dice:

what is the probability I see the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right? what is the probability I see one each of the numbers 6, 5, 4, 3, 2 and 1 on the dice regardless of order? Hint Have a look at the isin() function in the prob package. For example, to filter all rows containing 6, 1, 2 in that order, you could use isin(., c(6, 1, 2), ordered = TRUE)

```{r}
# Outcomes of rolling six dice
s_six_dice <- rolldie(6) 
```

```{r}
prob_6_5_4_3_2_1 <- nrow(1) / nrow(s_six_dice)
prob_6_5_4_3_2_1
```

The probability of seeing the numbers 6, 5, 4, 3, 2 and 1 in order on the dice, left to right is 1/46656 which R rounds to 0 My calculator tells me it is 0.00002143 - not at all likely then!

```{r}
# Hmm - can't get the code to work - but I can work it out manually
s_6_5_4_3_2_1_any_order <- s_six_dice %>% 
  isin(., c(1, 2, 3, 4, 5, 6))
```

```{r}
prob_6_5_4_3_2_1_any_order <- s_six_dice %>% 
  filter()
```

Manual working: Combinations of 6, 5, 4, 3, 2, 1 in any order = 6! = 6 x 5 x 4 x 3 x 2 x 1 = 720 We already know there are 46656 possible combinations p(of 6, 5, 4, 3, 2, 1 in any order) = 720/46656 = 0.0153421 - much more likely than in a specific order

## Question 6.

If I draw a single card from a 52-card deck, and it is black, what is the probability that its suit is spades?

Hints Investigate the cards() function in the prob package, if you'd like to try this in R (make a deck without jokers). You could think about adding your own black_card random variable to the sample space, which would be TRUE if the suit is 'Spade' or 'Club', and FALSE otherwise.

0.5 - as half of the black deck is clubs, half spades (or 13/26)
