---
title: "R Notebook"
output: html_notebook
---

```{r}
library(gtools)
library(tidyverse)
library(janitor)
library(skimr)
library(infer)
library(boot)
```

# 1 MVP

Now we'll go back to CI creation in the normal fashion. We'll take the ames data from the CIs lab earlier today and regard it now as a sample, we won't be drawing any smaller samples from within it. This is the usual situation in an analysis: you use all the data available to you!

## Task 1.

Load the data again, clean_names(), and re-familiarise yourself with it

```{r}
ames <- read_csv("data/ames.csv") %>%  clean_names()
```

## Task 2.

Investigate the distribution of lot_area. Is the distribution roughly normal? If not, what problems do you find?

```{r}
ames %>% 
  ggplot() +
  aes(x = lot_area) +
  geom_histogram(col = "white", bins = 25, fill = "violetred4")
```

A couple of outliers are skewing the data - making it appear right-skewed.

## Task 3.

Compute and visualise a bootstrap sampling distribution for the mean(lot_area) of the sold houses.

```{r}
ames_summary <- ames %>%
  summarise(
    num = n(),
    mean = mean(lot_area),
    sd = sd(lot_area)
  )
ames_summary
```

```{r}
# Create a bootstrap sample with 1000 means in samples of 100
bootstrap_resample_100 <- ames %>%
  rep_sample_n(size = 100, replace = TRUE, reps = 1000) %>%
  dplyr::summarise(
    mean_lot_area = mean(lot_area)
  )

bootstrap_resample_100
```

```{r}
bootstrap_resample_100 %>% 
  ggplot() +
  aes(x = mean_lot_area) +
  geom_histogram(col = "white", bins = 25, fill = "violetred4")
```

The bootstrap data appears to give a more normal distribution than the main data set, but outliers are still appearing and skewing the data.

## Task 4.

Use your bootstrap distribution to calculate a 95% CI for mean(lot_area), and visualise it on the distribution

```{r}
ci_95 <- bootstrap_resample_100 %>% 
  summarise(mean = mean(mean_lot_area),
            lower_bound = quantile(mean_lot_area, probs = 0.025),
            upper_bound = quantile(mean_lot_area, probs = 0.975))

ci_95
```

The mean lot area is 10143. Statistically 95% of the results will fall between 8989 and 12125.

## Task 5.

You would like to know the mean(lot_area) of the sold houses with higher confidence. Calculate the 99% CI for this variable (you can re-use your bootstrap distribution from above). Is it narrower or broader than the 95% CI? Does that make sense?

```{r}
ci_99 <- bootstrap_resample_100 %>% 
  summarise(mean = mean(mean_lot_area),
            lower_bound = quantile(mean_lot_area, probs = 0.005),
            upper_bound = quantile(mean_lot_area, probs = 0.995))

ci_99
```

The mean lot area remains at 10143. Statistically 99% of the results will fall between 8753 and 12607.

## Task 6.

Calculate the point estimate of the mean(lot_area)

```{r}
summary_bootstrap_resample_100 <- bootstrap_resample_100 %>%
  ungroup() %>%
  summarise(
    mean_mean_lot_area = mean(mean_lot_area)
  ) 

summary_bootstrap_resample_100
```

The mean (of the means) is 10143.

# 2 Extension

## Task 1.

Calculate a point estimate and 95% CI for the proportion of houses in the data built before 1920. Does the number of reps you use matter? \[Investigate reps from 200 up to 50000, memory of your laptop permitting\].

```{r}
# First mutate the data to add in pre/post 1920
ames <- ames %>% 
  mutate(before_1920 = if_else(year_built < 1921, "pre 1920", 
                               "1920 and after"),
  .after = year_built)
```

```{r}
# Create a bootstrap sample with 1000 means in samples of 100
bootstrap_resample_100 <- ames %>%
  rep_sample_n(size = 100, replace = TRUE, reps = 1000) %>%
  dplyr::summarise(
    mean_lot_area = mean(lot_area)
  )

bootstrap_resample_100
```

```{r}
bootstrap_resample <- ames %>%
  specify(response = before_1920, success = before_1920) %>%
  generate(reps = 500, type = "bootstrap") %>%
  calculate(stat = "prop")

head(bootstrap_resample)
```


Hmmm - clearly not understanding what is being asked for here - and not completed.

Tried with some code Emily had written - but don't think I was using it properly.

